---
title: "Betreuungsangebot für Kinder von 0-6"
subtitle: "Vergleich 2006/07-2016/17"
author: "Gerald Gartner"
date: "05-2015"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


```{r, echo=FALSE}
# CONFIG
user_name <- "qvvdata" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "kigaooe" # adapt!
package_date <- "2018-02-05" # date of the CRAN snapshot that
# the checkpoint package uses
```

## Notes

This report was generated on `r Sys.time()`.

...

### R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/rscript.zip). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `ìnput` will be processed and results will be written to `output`. 

### GitHub

The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 


### License

...

### `xyz.csv` 

...

## Preparations

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work, 
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}
```


### Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(magrittr) # pipes
library(stringr) # string manipulation
library(readxl) # excel
library(scales) # scales for ggplot2
library(jsonlite) # json
library(forcats) # easier factor handling,
library(lintr) # code linting
library(xlsx) #Excel
library(googlesheets) # googlesheets (replace with googlesheets4 asap)",
file = "manifest.R")
```

### Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("checkpoint",
                           username = "RevolutionAnalytics",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
checkpoint(snapshotDate = package_date,
           project = path_to_wd,
           verbose = T,
           scanForPackages = T,
           use.knitr = F)
rm(package_date)
```


### Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

### Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
knitr::read_chunk("scripts/theme_addendum.R")
source("./scripts/theme_addendum.R")
source("./scripts/BorderMan.R")
source("./scripts/01-preprocessing.R")
source('./scripts/geom_holygon.R')
cat(getwd())
#my_function(5)

```




```{r, echo=TRUE, message=FALSE, warning=FALSE}
options(scipen=999)


add_theme <- function(p) {
  p +
    labs(x = NULL, y = NULL) +
    theme(plot.background = element_blank(),
          panel.background = element_blank(),
           panel.grid = element_line(colour="transparent"),
           panel.grid.minor = element_line(colour="transparent"),
           panel.grid.major = element_line(colour="transparent"),
           legend.position="none",
           axis.title.x=element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank(),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
           panel.border=element_blank(),
           panel.spacing=unit(c(0,0,0,0), "null"),
           plot.margin=unit(c(0,0,0,0), "null")
  )
}

add_save <- function(p, filename) {
  ggsave(paste0("output/",filename,".png"), plot=p, width = 20, height = 10, units = "cm",
         bg = "transparent")
  ggsave(paste0("output/",filename,"_mobile.png"),plot=p, width = 20, height = 10, scale=0.25, units = "cm",
         bg = "transparent")}


```

Wo hat sich Öffnungszeit positiv/negativ entwickelt?
Welche Stadt/Land-Unterschiede bzw. nach Bundesländern gibt es?
Wie haben sich Schließtage entwickelt? Welche Stadt/Land-Unterschiede gibt es?

Auf welche Erhalter geht der Ausbau zurück? Sind private/öffenltiche dafür verantwortlich?
In welchen Gemeinden gibt es im Vergleich zu 2006 erstmals eine Betreuungseinrichtung?

Ist frühkindliche Betreuung ausgebaut worden? Wenn ja, wo? Wie hat sich Betreuungsquote verändert?

Bei altersgemischten Einrichtungen werden die GEwichte für 3 bis 6-Jährige miteinbezogen, nicht für alle anderen Altersgruppen. 26209 Kinder im betreffenden Alter sind in einer altersgemischten Einrichtung. Kindergartenkinder zählt die Kindertagesheimstatistik 221.490. 

```{r}
# print(length(which(std$diffmin <0)))
# print(length(which(std$diffmin ==0)))
# print(length(which(std$diffmin >0)))
# 
# print(length(which(std$diffmin >0 && std$typgmd==1)))
# print(length(which(std$diffmin ==0 && std$typgmd==1)))
#print(length(which(std$gemgew16 >480))

#In wie vielen hat es sich am Land/Stadt verlängert/verkürzt?
countup<- std %>%
        #filter(diffmin<0) %>% 
        group_by(typgmdtxt, diffmintxt) %>% 
        count() %>%
                summarise(sum = sum(n))


cat(getwd())
write.csv(countup, file ="output/datawrapper/laengerkurzer.csv")

std$gemgew16_over8h = std$gemgew16>480
table(std$gemgew16_over8h)




```


```{r}
#Visualisierungen

  # labs(title = "Schlagzeile", 
  #      x = "Vertikale Achse", 
  #      y = "horizontale Achse",
  #      #subtitle ="Funktioniert aus unerklärlichen Gründen noch nicht", 
  #      caption = "Quelle")+
  # theme_addendum()

#Exportieren für Illustrator funktioniert nicht über ggsave (Schriften), sondern besser über Export des Plot-Viewers#

# 
#needs(magrittr, maptools, Cairo, ggmap, maps, rgdal, raster, viridis, readxl, googlesheets, gridExtra)
library(rgdal)
library(rgeos)
needs(tidyverse, magrittr)
gde_18 <- readOGR("input/geo/2018/gemeinden m bezirke 2018.geojson","OGRGeoJSON")

gde_18@data$BL = substring(gde_18@data$GKZ,0,1)

#region <- spChFIDs(region, row.names(region))
bundeslaendergrenzen <- gUnaryUnion(gde_18, id = gde_18@data$BL)



map_data_fortified <- fortify(gde_18, region = "GKZ") %>% 
  mutate(id = as.numeric(id))

map_data <- std %>% left_join(map_data_fortified, by = c("gkz" = "id")) 
map_data$klasse <- cut(map_data$diffmin, c(-220,-35,-5, 5,18,70,330),  labels = c("viel kürzer", "kürzer", "gleich" ,"etwas länger", "länger","viel länger"))

quantile(std$diffmin, na.rm = TRUE, probs = seq(0, 1, 0.01))
#  0%         10%         20%         30%         40%         50%         60%         70%         80%         90%        100% 
# -216.571429  -36.000000   -9.150000    0.000000    1.853223   18.000000   36.000000   58.620360   85.300961  141.000000  326.400000 

map_data$diffminmap <- map_data$diffmin
map_data$diffminmap[map_data$diffmin <= -60] <- -60
map_data$diffminmap[map_data$diffmin >= 203] <- 203

#VISUALISIERUNG ZEICHNEN
p7 <- ggplot() +
    # municipality polygons
    geom_polygon(data = map_data, aes(x = long, 
                                      y = lat, 
                                      group = group, 
                                      fill=diffminmap)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.001) +
    coord_map(projection = "mercator") +
    scale_fill_gradientn(breaks=c(min(map_data$diffminmap),0,max(map_data$diffminmap)),
                       colors=c('#ba2b58','#f0edf1','#84a07c')) +
    geom_holygon(data = bundeslaendergrenzen,
               fill = "transparent", color="white",
               size=0.5, aes(x=long,y=lat,group=group)) +
    labs(x = NULL, 
         y = NULL, 
         title = "Kinderbetreuungszeitenvergleich", 
         #subtitle = "Vergleich der Erwerbsquoten je Geschlecht", 
         caption = "Quelle: Statistik Austria, BEV.")+
  guides(fill=guide_legend(nrow=6,byrow=TRUE))
plot(p7)

ggsave("p7.pdf", width = 15, height = 60, units = "cm")
add_save(add_theme(p7 + labs(title=NULL, caption=NULL)),'kinderbetreuungszeiten_vergleich')

#source("https://data.addendum.org/gfx/fonts/theme_addendum.R", font_import())
```

```{r}


p <- ggplot() +
    # municipality polygons
    geom_polygon(data = map_data, aes(x = long, 
                                      y = lat, 
                                      group = group, 
                                      fill=gemgew16)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.001) +
    coord_map(projection = "mercator") +
    scale_fill_gradientn(breaks=c(0,min((map_data %>% filter(gemgew16>0))$gemgew16), max(map_data$gemgew16)),
                       colors=c('white','#DCD7DE','#40234b')) +
    geom_holygon(data = bundeslaendergrenzen,
               fill = "transparent", color="white",
               size=0.5, aes(x=long,y=lat,group=group)) +
    labs(x = NULL, 
         y = NULL, 
         title = "Gute Kinderbetreuung, schlechte Kinderbetreuung", 
         #subtitle = "Vergleich der Erwerbsquoten je Geschlecht", 
         caption = "Quelle: Statistik Austria, BEV.")+
  guides(fill=guide_legend(nrow=6,byrow=TRUE))
plot(p)
add_save(add_theme(p + labs(title=NULL, caption=NULL)),'kinderbetreuungszeiten_2016')
```

```{r}
p <- ggplot() +
    # municipality polygons
    geom_polygon(data = map_data, aes(x = long, 
                                      y = lat, 
                                      group = group, 
                                      fill=gemgew06)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.001) +
    coord_map(projection = "mercator") +
    scale_fill_gradientn(breaks=c(0,min((map_data %>% filter(gemgew06>0))$gemgew06), max(map_data$gemgew06)),
                       colors=c('white','#d9d9d9','#777777')) +
    geom_holygon(data = bundeslaendergrenzen,
               fill = "transparent", color="white",
               size=0.5, aes(x=long,y=lat,group=group)) +
    labs(x = NULL, 
         y = NULL, 
         title = "Kinderbetreuungszeitenvergleich", 
         #subtitle = "Vergleich der Erwerbsquoten je Geschlecht", 
         caption = "Quelle: Statistik Austria, BEV.")+
  guides(fill=guide_legend(nrow=6,byrow=TRUE))
plot(p)
add_save(add_theme(p + labs(title=NULL, caption=NULL)),'kinderbetreuungszeiten_2006')
```




```{r}

map_data$gemgew16_over8h = map_data$gemgew16>480

p <- ggplot() +
    # municipality polygons
    geom_polygon(data = map_data, aes(x = long, 
                                      y = lat, 
                                      group = group, 
                                      fill=gemgew16_over8h)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.001) +
    coord_map(projection = "mercator") +
    scale_fill_manual(breaks=c(TRUE, FALSE), values=c('TRUE'='#CDCC32','FALSE'='white')) +
#    scale_fill_gradientn(breaks=c(0,min((map_data %>% filter(gemgew06>0))$gemgew06), max(map_data$gemgew06)),
#                       colors=c('white','#d9d9d9','#777777')) +
    geom_holygon(data = bundeslaendergrenzen,
               fill = "transparent", color="white",
               size=0.5, aes(x=long,y=lat,group=group)) +
    labs(x = NULL, 
         y = NULL, 
         title = "Kinderbetreuungszeitenvergleich", 
         #subtitle = "Vergleich der Erwerbsquoten je Geschlecht", 
         caption = "Quelle: Statistik Austria, BEV.")+
  guides(fill=guide_legend(nrow=6,byrow=TRUE))
plot(p)
add_save(add_theme(p + labs(title=NULL, caption=NULL)),'kinderbetreuungszeiten_2016_ueber8h')
```


## Linting

The code in this RMarkdown is listed with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd")
# if you have additional scripts and want them to be linted too, add them here
#lintr::lint("scripts/my_script.R")
```


