---
title: "Betreuungsangebot für Kinder von 0-6"
subtitle: "Vergleich 2006/07-2016/17"
author: "Gerald Gartner"
date: "05-2015"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


```{r, echo=FALSE}
# CONFIG
user_name <- "qvvdata" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "kigaooe" # adapt!
package_date <- "2018-02-05" # date of the CRAN snapshot that
# the checkpoint package uses
```

## Notes

This report was generated on `r Sys.time()`.

...

### R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/rscript.zip). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `ìnput` will be processed and results will be written to `output`. 

### GitHub

The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 


### License

...

### `xyz.csv` 

...

## Preparations

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work, 
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}
```


### Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(magrittr) # pipes
library(stringr) # string manipulation
library(readxl) # excel
library(scales) # scales for ggplot2
library(jsonlite) # json
library(forcats) # easier factor handling,
library(lintr) # code linting
library(xlsx) #Excel
library(googlesheets) # googlesheets (replace with googlesheets4 asap)",
file = "manifest.R")
```

### Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("checkpoint",
                           username = "RevolutionAnalytics",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
checkpoint(snapshotDate = package_date,
           project = path_to_wd,
           verbose = T,
           scanForPackages = T,
           use.knitr = F)
rm(package_date)
```


### Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

### Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
knitr::read_chunk("scripts/theme_addendum.R")
source("./scripts/theme_addendum.R")
source("./scripts/BorderMan.R")
source("./scripts/01-preprocessing.R")
cat(getwd())
#my_function(5)

```




```{r, echo=TRUE, message=FALSE, warning=FALSE}
options(scipen=999)

#Analyse

```

Wo hat sich Öffnungszeit positiv/negativ entwickelt?
Welche Stadt/Land-Unterschiede bzw. nach Bundesländern gibt es?
Wie haben sich Schließtage entwickelt? Welche Stadt/Land-Unterschiede gibt es?

Auf welche Erhalter geht der Ausbau zurück? Sind private/öffenltiche dafür verantwortlich?
In welchen Gemeinden gibt es im Vergleich zu 2006 erstmals eine Betreuungseinrichtung?

Ist frühkindliche Betreuung ausgebaut worden? Wenn ja, wo? Wie hat sich Betreuungsquote verändert?


```{r}
#Visualisierungen

  # labs(title = "Schlagzeile", 
  #      x = "Vertikale Achse", 
  #      y = "horizontale Achse",
  #      #subtitle ="Funktioniert aus unerklärlichen Gründen noch nicht", 
  #      caption = "Quelle")+
  # theme_addendum()

#Exportieren für Illustrator funktioniert nicht über ggsave (Schriften), sondern besser über Export des Plot-Viewers#

# 
#needs(magrittr, maptools, Cairo, ggmap, maps, rgdal, raster, viridis, readxl, googlesheets, gridExtra)
library(rgdal)
needs(tidyverse, magrittr)
gde_18 <- readOGR("input/geo/2018/gemeinden_w_bez.shp", layer = "gemeinden_w_bez")



map_data_fortified <- fortify(gde_18, region = "GKZ") %>% 
  mutate(id = as.numeric(id))

map_data <- std %>% left_join(map_data_fortified, by = c("gkz" = "id")) 
map_data$klasse <- cut(map_data$diffmin, c(-220,-75,-5, 5,18,70,330),  labels = c("viel kürzer", "kürzer", "gleich" ,"etwas länger", "länger","viel länger"))

quantile(std$gemgew16, na.rm = TRUE, probs = seq(0, 1, 0.1))
#  0%         10%         20%         30%         40%         50%         60%         70%         80%         90%        100% 
# -216.571429  -36.000000   -9.150000    0.000000    1.853223   18.000000   36.000000   58.620360   85.300961  141.000000  326.400000 

map_data$diffminmap <- map_data$gemgew16
map_data$diffminmap[map_data$gemgew16 >= 600] <- 600
map_data$diffminmap[map_data$gemgew16 <= 360] <- 360

#VISUALISIERUNG ZEICHNEN
p7 <- ggplot() +
    # municipality polygons
    geom_polygon(data = map_data, aes(x = long, 
                                      y = lat, 
                                      group = group, 
                                      fill=diffminmap)) +
    # municipality outline
    geom_path(data = map_data, aes(x = long, 
                                   y = lat, 
                                   group = group), 
              color = "white", size = 0.001) +
    coord_equal() +
    coord_map(projection = "mercator")+
# scale_fill_manual(values=c("viel länger" = "#36513E",
#                           "länger" = "#52705B",
#                            "etwas länger" = "#869B8C",
#                           "gleich" = "#A8A8A8",
#                             "kürzer" = "#A89BAD", 
#                               "viel kürzer" = "#40234b"))+
  scale_fill_gradient2(low = "#ba2b58", midpoint = 500, mid = "#f0edf1", high = "#84a07c")+
    labs(x = NULL, 
         y = NULL, 
         title = "Kinderbetreuungszeitenvergleich", 
         #subtitle = "Vergleich der Erwerbsquoten je Geschlecht", 
         caption = "Quelle: Statistik Austria, BEV.")+
   theme(legend.position="top",
      legend.text = element_text(size=8),
      axis.line=element_blank(),
      strip.background = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())+
  guides(fill=guide_legend(nrow=6,byrow=TRUE))
plot(p7)
ggsave("p7.pdf", width = 15, height = 60, units = "cm")

#source("https://data.addendum.org/gfx/fonts/theme_addendum.R", font_import())
```

## Linting

The code in this RMarkdown is listed with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd")
# if you have additional scripts and want them to be linted too, add them here
#lintr::lint("scripts/my_script.R")
```


